%% ============================================================
%%  BOOK 2 — COMPLETE OUTLINE
%%  Topographical Orthogonal Generative Theory:
%%  Applications, Verification, and the Foundations of All Domains
%%  Pablo Nogueira Grossi · G6 LLC · Newark NJ · 2026
%% ============================================================

\documentclass[12pt, oneside]{book}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{booktabs}
\usepackage{enumitem}
\geometry{paperwidth=5.5in, paperheight=8.5in,
  top=0.875in, bottom=0.875in, inner=0.875in, outer=0.625in}

\begin{document}
\thispagestyle{empty}

\begin{center}
{\LARGE\bfseries Book 2 Outline}\\[0.5em]
{\large Topographical Orthogonal Generative Theory}\\[0.3em]
{\large Applications, Verification, and the Foundations of All Domains}\\[1em]
{\normalsize Pablo Nogueira Grossi \textperiodcentered\ G6 LLC
\textperiodcentered\ Newark NJ \textperiodcentered\ 2026}
\end{center}

\vspace{0.5cm}
\hrule
\vspace{0.5cm}

%% ── FRONT MATTER ─────────────────────────────────────────────

\noindent\textbf{Front Matter}
\begin{itemize}[leftmargin=2em]
\item Half title
\item Title page
\item Copyright · ISBN · G6 LLC · Zenodo DOIs
\item Dedication --- to Vic, Giulia, Alice, Sarah, David
\item Letter to the Holy Father (carried from Book 1)
\item Preface \textit{[written]}
\item Table of Contents
\end{itemize}

\vspace{0.5cm}
\hrule
\vspace{0.5cm}

%% ── PART I ───────────────────────────────────────────────────

\noindent\textbf{Part I: The Formal Verification}\\
\textit{The mathematics made machine-checkable.}

\medskip
\noindent\textbf{Chapter 1: The Translation} \textit{[written]}\\
The governing idea. What Book 1 established. What this book does.
The scale problem. The Grossi Generative Science. How to read this book.

\medskip
\noindent\textbf{Chapter 2: AXLE --- The Proof Engine}\\
What AXLE is and why it exists. The Lean 4 / Mathlib setup.
The operator chain as Lean types: \texttt{CompressionOp},
\texttt{CurvatureOp}, \texttt{FoldOp}, \texttt{UnfoldOp},
\texttt{GenerativeOp}. The dm3 canonical invariants as proved
theorems. The G6 Crystal invariants: \texttt{crystal\_aspect\_ratio},
\texttt{g6\_equals\_schumann}. The commit log as a mathematical
record. How to clone and build.

\medskip
\noindent\textbf{Chapter 3: The Regeneration Hierarchy}\\
From $\mathbb{N}$ to the ordinals. \texttt{nextLevel} by construction.
\texttt{OrdinalRegenerationLevel}. The club filter:
\texttt{IsUnboundedBelow}, \texttt{IsOmegaClosedBelow},
\texttt{IsClubBelow}, \texttt{IsStationaryBelow}.
\texttt{closurePoints\_stationary} for regular uncountable cardinals.
The Volume IV master theorem: \texttt{regeneration\_hierarchy\_mahlo}.
What remains: Issue \#6, the full proof without regularity hypothesis.

\medskip
\noindent\textbf{Chapter 4: Eight Things That Broke}
\textit{[written]}\\
What building AXLE revealed about large language models.
Eight observations, eight fixes. The independent researcher gap.

\vspace{0.5cm}
\hrule
\vspace{0.5cm}

%% ── PART II ──────────────────────────────────────────────────

\noindent\textbf{Part II: The Applications}\\
\textit{The operator sequence instantiated across domains.}

\medskip
\noindent\textbf{Chapter 5: The G6 Crystal}\\
Full architectural derivation from dm3 invariants.
Hexagonal cross-section: isoperimetric optimum proved.
Hexagrid structural system: peer-reviewed performance verified.
Schumann $n=4$ resonance coupling: Arnold tongue $\mathcal{A}_{4:1}$,
noise tolerance $\tau \cdot \varepsilon_0 = 2/3$.
Tropopause as limit cycle $\Gamma$.
The four Civilizational Theorems: Mars gravity scaling,
pressure vessel stability, dust storm resonance tolerance,
material self-sufficiency. The falsifiable prediction:
scale model at 33.5\,Hz.

\medskip
\noindent\textbf{Chapter 6: The Fruit Fly Connectome}\\
AXLE as toy machine. The connectome as a graph instantiation
of $G = U \circ F \circ K \circ C$. Loading the FlyWire data.
$C$ as synaptic pruning. $K$ as clustering-coefficient reweighting.
$F$ as fold-node collapse (Whitney $A_1$). $U$ as PageRank
attractor selection. The dm3 metric $\tau = \sqrt{c/\kappa}$
computed on neural clusters. Comparison to canonical
$(T^*, \mu_{\max}, \tau) = (2\pi, -2, 2)$. Arnold tongue check.
Three predictions for experimental neuroscience.

\medskip
\noindent\textbf{Chapter 7: Plasma Instabilities}\\
Magnetic reconnection as a generative transition.
$C$: flux compression in the Harris current sheet.
$K$: current sheet thinning toward $\kappa^*$.
$F$: X-point onset, rank-1 Jacobian loss.
$U$: Taylor relaxation to minimum-energy state.
The reconnection rate $\gamma$ as $\mu_{\max}$.
The Sweet-Parker length as $\varepsilon_0$.
Prediction: reconnection onset threshold from dm3 constants.

\medskip
\noindent\textbf{Chapter 8: String Theory Mapping}\\
NS-NS and RR sector operators mapped to $C \to K \to F \to U$.
Compactification as $C$. Moduli stabilization as $K$.
Conifold transition as $F$. Flux stabilization (GVW) as $U$.
The axio-dilaton $\tau_\text{IIB}$ as embodiment threshold.
The perturbative regime boundary as $\varepsilon_0 = 1/3$.
What the framework predicts about vacuum selection.

\medskip
\noindent\textbf{Chapter 9: Martian Colony Architecture}\\
Closed-loop life support as operator chain.
$C$: metabolic compression to minimal viability set.
$K$: physiological and engineering stress accumulation.
$F$: system fold --- atmosphere breach, crop failure, radiation.
$U$: redundant system activation, biological adaptation.
Allostatic unification law applied to coupled subsystems.
The G6 Crystal scaled to Mars gravity ($\gamma^{-1} \approx 2.64$).
Mars height 39,808\,m, well within Martian troposphere.
ISRU material self-sufficiency: six layers seed six modules.

\medskip
\noindent\textbf{Chapter 10: Finance and Economic Systems}\\
Price trajectories as generative transitions.
Market fold events: Black Monday, the 2008 crisis, flash crashes.
$C$: reduction of market degrees of freedom to dominant modes.
$K$: volatility accumulation toward critical threshold.
$F$: liquidity bifurcation, rank-1 order book collapse.
$U$: price stabilization on new branch (circuit breaker, bailout).
$\tau$ as systemic risk threshold. $\varepsilon_0 = 1/3$ as
the margin before structural failure of the market manifold.
Prediction: pre-fold signatures detectable from dm3 invariants.

\medskip
\noindent\textbf{Chapter 11: Language, Meaning, and Compression}\\
Semantic compression as operator $C$.
Metaphor as fold: rank-1 collapse of meaning.
Interpretation as unfold: recovery of stable branch.
The Maggid tradition as a historical instance of the operator:
transmission is compression, reception is unfold.
The Torah as a compressed encoding of structure that the
mathematical framework now translates.
Not mysticism: archaeology.

\vspace{0.5cm}
\hrule
\vspace{0.5cm}

%% ── PART III ─────────────────────────────────────────────────

\noindent\textbf{Part III: The Foundations}\\
\textit{What the framework is, at depth.}

\medskip
\noindent\textbf{Chapter 12: The Separation Theorem}\\
Theorem 12.2 of TOGT Volume I (cited in the crystal paper):
$\mathrm{Tr}(M^6) \neq 33$ for any stable $n < 33$ matrix
representation. Therefore $g^6 = 33$ is a topological invariant
of the six-iterate operator algebra, not an algebraic coincidence.
Full proof. Implications for the Schumann coupling: the integer
33 cannot be produced by lower-dimensional representations.
It lives at the topological layer.

\medskip
\noindent\textbf{Chapter 13: Large Cardinals and the Operator Algebra}\\
The regeneration hierarchy as an ordinal structure.
Mahlo cardinals: every club contains a regular cardinal.
The TOGT Mahlo condition: every club contains a regeneration
closure point. The hyper-Mahlo claim: the fixed points are
stationary. Issue \#6: proof without regularity hypothesis.
What inaccessible cardinals mean for the operator algebra.
The connection to Woodin cardinals and the ultimate $L$ program.
What remains open.

\medskip
\noindent\textbf{Chapter 14: The Volume IV Program}\\
Honor. Lineage. Restoration. The three nested circles:
outer (published mathematics for all), middle (specific
mathematics for specific people, transmitted not published),
inner (higher mathematics, only for the author).
The Templar topology: suppression as $B_3$ (collapse boundary),
the Chinon Parchment as the private absolution before the
public condemnation. The Order of Christ as $U$: stabilization
on a new branch after the fold.
The mathematical structure of historical justice.

\vspace{0.5cm}
\hrule
\vspace{0.5cm}

%% ── BACK MATTER ──────────────────────────────────────────────

\noindent\textbf{Back Matter}
\begin{itemize}[leftmargin=2em]
\item Appendix A: AXLE Repository Reference ---
      full theorem list, proof status, issue log
\item Appendix B: The Crystal Paper ---
      \textit{The G6 Crystal: A dm3-Derived Architectural Form}
      (Zenodo 10.5281/zenodo.19162013), reprinted in full
\item Appendix C: Canonical Values Reference Card ---
      all locked constants, formulas, and their derivations
\item Bibliography
\item About the Author
\item Manifesto --- with Pope's blessing
\item Colophon
\end{itemize}

\vspace{0.5cm}
\hrule
\vspace{0.5cm}

%% ── CHAPTER COUNT AND STATUS ─────────────────────────────────

\noindent\textbf{Status as of March 22, 2026}

\medskip
\begin{tabular}{llll}
\toprule
\textbf{Chapter} & \textbf{Title} & \textbf{Status} & \textbf{Notes} \\
\midrule
Preface & The Translation & Written & Complete \\
Ch.\ 1  & The Translation & Written & Complete \\
Ch.\ 2  & AXLE & Drafted & Needs polish \\
Ch.\ 3  & Regeneration Hierarchy & Drafted & Issue \#6 pending \\
Ch.\ 4  & Eight Things That Broke & Written & Complete \\
Ch.\ 5  & G6 Crystal & Written & Crystal paper complete \\
Ch.\ 6  & Fruit Fly Connectome & Scaffolded & AXLE sims ready \\
Ch.\ 7  & Plasma & Outlined & \\
Ch.\ 8  & String Theory & Outlined & \\
Ch.\ 9  & Martian Colony & Outlined & Crystal paper §6 \\
Ch.\ 10 & Finance & Outlined & \\
Ch.\ 11 & Language \& Meaning & Outlined & \\
Ch.\ 12 & Separation Theorem & Stated & Proof needed \\
Ch.\ 13 & Large Cardinals & Drafted & Issue \#6 \\
Ch.\ 14 & Volume IV Program & Outlined & \\
\bottomrule
\end{tabular}

\bigskip
\noindent\textbf{Estimated page count:} 280--320 pages.\\
\textbf{ISBN:} to be assigned (G6 LLC).\\
\textbf{Target:} IngramSpark, 60lb cream, matte laminate.

\vspace{0.5cm}

\begin{center}
\itshape
C \;$\to$\; K \;$\to$\; F \;$\to$\; U \;$\to$\; $\infty$\\[0.3em]
\normalfont\small G6 LLC \textperiodcentered\ Newark NJ
\textperiodcentered\ 2026
\end{center}

\end{document}
