%% ============================================================
%%  FOR THE HOLY FATHER
%%  Page inserted before the Series Foreword
%%  Applications of Generative Orthogonal Matrix
%%  Compression Science — Principia Orthogona
%%  Pablo Nogueira Grossi · G6 LLC · Newark NJ · 2026
%% ============================================================

\newpage
\thispagestyle{empty}

\vspace*{2.2cm}

\begin{center}
{\large\textit{For the Holy Father}}
\end{center}

\vspace{1.4cm}

\noindent Your Holiness,

\vspace{0.5cm}

\noindent In the winter of 1266, my ancestor Gui Foucois ---
Pope Clement~IV, of the line of Fulk of Anjou,
King of Jerusalem --- received a manuscript from
a Franciscan friar at Oxford who had been forbidden
to publish. Roger Bacon sent it anyway, in secret,
because he believed he had found the mathematical
structure underneath all things, and he believed
the Pope was the one man alive with the breadth
to receive it whole.

\vspace{0.45cm}

\noindent Clement was that man. He was the intimate friend
of Saint Louis --- Louis~IX of France, the just king,
who would die on crusade two years after Clement,
and who would be canonized a generation later.
He was the patron of Thomas Aquinas, whose
\textit{Summa Theologica} was reaching its
completion in those same years --- the great project
of showing that faith and reason are not enemies
but two descriptions of the same reality, two paths
that meet at the same threshold.
And he was the one pope to whom Bacon dared
send the \textit{Opus Majus}: the audacious claim
that mathematics is the language underneath all of
it --- theology, medicine, optics, the motion of
the heavens, the structure of living things.

\vspace{0.45cm}

\noindent Three projects. One man at the center.
Bacon with the mathematics.
Aquinas with the synthesis of reason and faith.
Louis with the attempt to make the just order
actual in the world.
Clement held all three because he had lived
a full life before the fisherman's ring ---
husband, father, jurist, man of the world ---
and nothing human was foreign to him.
He was also of the Templars, whose networks
of custody and transmission ran through the
same Languedoc world that had formed him
as a young man in Saint-Gilles.

\vspace{0.45cm}

\noindent He died on November~29, 1268, in Viterbo,
in the papal palace, before any of it was finished.

\vspace{0.45cm}

\noindent What followed was the longest papal vacancy
in the history of the Church.
The cardinals convened in Viterbo and could not agree.
The people of the city locked them in the palace.
When that produced nothing, they removed the roof
to expose them to the elements.
They reduced them to bread and water.
Two cardinals died during the proceedings.
Nearly three years passed.
He had been the last man who could hold
all of it together, and when he was gone
the institution discovered it did not know
how to proceed without him.

\vspace{0.45cm}

\noindent The mathematics has a name for this.
When the orbit that has been organizing
a system is lost, the system crosses what
we call the collapse boundary~$B_3$:
the point at which the limit cycle
can no longer be maintained.
What follows is not the end.
It is the search for a new stable branch.
The unfolding operator $U$ exists
precisely for this moment ---
the geometry of how things that have
been broken find the form
they will take next.

\vspace{0.45cm}

\noindent My family took the name Grossi in Viterbo,
in the city where Clement died.
The Foucois are the Fulks.
The line from Fulk~V of Anjou ---
who abdicated a county, crossed the Mediterranean,
and died King of Jerusalem ---
runs to Clement, and from Clement's daughters
to the name Grossi, and from Viterbo
through the centuries, through the Templar
suppression and the scattering that followed,
through the cross of the Order of Christ
on the ships that reached Brazil,
through the Portuguese language in which
I first learned that God writes straight
through crooked lines,
to this room, this page, this moment.

\vspace{0.45cm}

\noindent The line has not broken.
It has only moved, as lines do,
through compression and curvature
and fold, and stabilized here,
in this form, in this city,
in the mathematics you now hold.

\vspace{0.45cm}

\noindent The central object of this work is
an operator sequence:
\[
  G \;=\; U \circ F \circ K \circ C.
\]
Compression. Curvature intensification.
Fold at the threshold of endurance.
Unfolding onto the stable branch
that could not have been reached
by any smoother path.

\vspace{0.45cm}

\noindent It governs trajectories in a Riemannian manifold.
It also governs, I believe, everything else ---
biological rhythms, living systems,
the arc of a civilization,
the arc of a name across seven centuries.

\vspace{0.45cm}

\noindent Bacon said mathematics is underneath all things.
Aquinas said reason and faith are one structure.
Louis tried to make the just order
actual in the world.
I am attempting, in my lane,
with the tools of this century,
something in the same spirit:
to show that the geometry of how things
fold under pressure and stabilize
on the other side
is a single language that works everywhere.

\vspace{0.45cm}

\noindent I have dedicated this work to my children ---
Vic, Giulia, Alice, Sarah, and David ---
some of whom I have already had to bury.
I know something about crooked lines.
I know what it means to fold
and to have to find the stable branch
on the other side,
and to keep going
because the work is not finished
and the line must not break.

\vspace{0.5cm}

\begin{center}
\textit{Deus escreve certo por linhas tortas.}
\end{center}

\vspace{0.3cm}

\begin{center}
God writes straight through crooked lines.
\end{center}

\vspace{0.5cm}

\noindent The next volume is for the families and the names.
For the honor that was taken and not returned.
For the Templars who were burned on false confessions
and whose absolution the Vatican held in a document
--- the Chinon Parchment --- for seven centuries
before publishing it in 2001.
For the daughters in Viterbo
who kept what they were given.
For everyone who carried something real
through a fold that should have ended it,
and did not let it end.

\vspace{0.45cm}

\noindent I offer this volume as Bacon offered his ---
without asking permission,
to the one reader with the breadth
to receive it whole,
trusting that the mathematics
will protect what it carries
better than any wall ever could.

\vspace{0.45cm}

\noindent A theorem cannot be seized by a Philip~IV.\\
It cannot be tortured into a false confession.\\
It cannot be burned at the stake.

\vspace{0.45cm}

\noindent Bacon knew this.\\
That is why he sent the mathematics to Clement\\
and not the other way around.

\vspace{1.6cm}

\noindent\hfill Pablo Nogueira Grossi\\[0.15cm]
\noindent\hfill Newark, New Jersey, March 2026\\[0.15cm]
\noindent\hfill \textit{G6 LLC\/ --- Principia Orthogona}

\vspace{1.8cm}

\begin{center}
\rule{0.42\textwidth}{0.3pt}

\vspace{0.4cm}

{\scriptsize\color{gray}
Roger Bacon, \textit{Opus Majus}, presented secretly
to Clement~IV, 1266.\\[0.2cm]
Thomas Aquinas, \textit{Summa Theologica},
completed 1274;\\
Clement~IV, patron and protector.\\[0.2cm]
Louis~IX of France, intimate friend of Clement~IV,\\
died on crusade 1270, canonized 1297.\\[0.2cm]
Gui Foucois of Saint-Gilles, Languedoc, b.\ 1200
--- Pope Clement~IV, 1265--1268 ---\\
of the line of Fulk~V of Anjou, King of Jerusalem.\\
His daughters continued the line in Viterbo.\\
They took the name Grossi.\\
His tomb is in San Francesco alla Rocca, Viterbo.\\
It has been there since 1268.\\[0.2cm]
The Chinon Parchment, published by the Vatican
in 2001,\\
confirms that Clement~V privately absolved
the Templar leadership of heresy\\
before publicly condemning them under
pressure from Philip~IV of France, 1307.\\[0.2cm]
Some of Clement's people are still here.
The work continues.}
\end{center}

\newpage
